Reanalysis Step 7: Sample Comparison Table

### 2025-09-04 reanalysis-7-missingness_patterns.qmd
data comparison

library(tableone)
library(knitr)
library(dplyr)

@@

@@

# Variables in original data (with proper capitalization)
vars_categorical_original <- c("gender_cat", "race_ethnic_cat", "insurance_cat", "age_cat",
                               "ever_lapse_binary", 
                               "glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat",
                               "catsurg_before_hitplus_cat",
                               "anti_VEGF", "PRP_flag", "focal_laser_flag", "other_inject",
                               "outcome_va_vi_binary")

# Additional variables in new data only
vars_categorical_new_only <- c("dr_severity")

@@

```{r create-table}
# Process original data (with original variable names)
original_summary <- process_dataset(original_complete, "Original", vars_categorical_original)

# Process submitted (included) data
submitted_summary <- process_dataset(submitted_data, "Submitted", vars_categorical_new)

# Process excluded data
excluded_summary <- process_dataset(excluded_data, "Excluded", vars_categorical_new)

@@

@@

# Create formatted table for categorical variables
format_categorical_table <- function(original, submitted, excluded, imputed) {
  vars <- unique(c(original$variable, submitted$variable, excluded$variable, imputed$variable))

@@

    # Original
    orig_row <- original[original$variable == v, ]
    if (nrow(orig_row) > 0 && !is.na(orig_row$n)) {
      table_df$Original[i] <- sprintf("%.0f (%.1f%%)", orig_row$n, orig_row$pct)
    }

    # Submitted
    sub_row <- submitted[submitted$variable == v, ]
    if (nrow(sub_row) > 0 && !is.na(sub_row$n)) {
      table_df$Submitted[i] <- sprintf("%.0f (%.1f%%)", sub_row$n, sub_row$pct)
    }

    # Excluded
    exc_row <- excluded[excluded$variable == v, ]
    if (nrow(exc_row) > 0 && !is.na(exc_row$n)) {
      table_df$Excluded[i] <- sprintf("%.0f (%.1f%%)", exc_row$n, exc_row$pct)
    }

    # Imputed
    imp_row <- imputed[imputed$variable == v, ]
    if (nrow(imp_row) > 0 && !is.na(imp_row$n)) {
      table_df$Imputed[i] <- sprintf("%.1f (%.1f%%)", imp_row$n, imp_row$pct)
    }
  }

>>>>

library(tableone)
library(knitr)
library(dplyr)
library(tidyr)

@@

# Create dr_severity for original data from No_DR/NPDR/PDR
original_complete <- original_complete |>
  mutate(
    dr_severity = case_when(
      No_DR == TRUE ~ "No_DR",
      NPDR %in% c("Mild", "Moderate", "Severe") ~ "NPDR",
      PDR == "Present" ~ "PDR",
      TRUE ~ NA_character_
    ),
    dr_severity = factor(dr_severity, levels = c("No_DR", "NPDR", "PDR"))
  )

@@

# Variables in original data (with proper capitalization)
vars_categorical_original <- c("gender_cat", "race_ethnic_cat", "insurance_cat", "age_cat",
                               "dr_severity", "ever_lapse_binary", 
                               "glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat",
                               "catsurg_before_hitplus_cat",
                               "anti_VEGF", "PRP_flag", "focal_laser_flag", "other_inject",
                               "outcome_va_vi_binary")

# Additional variables (none now that dr_severity is in both)
vars_categorical_new_only <- c()

@@

```{r create-table}
# Process original data (with original variable names)
original_summary <- process_dataset(original_complete, "Original", vars_categorical_original)
original_summary$categorical$variable_std <- original_summary$categorical$variable

# Process submitted (included) data
submitted_summary <- process_dataset(submitted_data, "Submitted", vars_categorical_new)
submitted_summary$categorical$variable_std <- submitted_summary$categorical$variable

# Process excluded data
excluded_summary <- process_dataset(excluded_data, "Excluded", vars_categorical_new)
excluded_summary$categorical$variable_std <- excluded_summary$categorical$variable

@@

# Add variable_std column to average_imputed_categorical for consistency
average_imputed_categorical$variable_std <- average_imputed_categorical$variable

@@

# Create formatted table for categorical variables
format_categorical_table <- function(original, submitted, excluded, imputed) {
  # Get unique base variable names (without the level suffixes)
  get_base_var <- function(x) sub("_[^_]+$", "", x)

  # Collect all variables
  all_vars <- unique(c(original$variable, submitted$variable, excluded$variable, imputed$variable))

  # For glaucoma and otherretina, we need to handle different naming conventions
  # Original has _FALSE/_TRUE, new data has _0/_1
  standardize_var <- function(v) {
    if (grepl("glaucoma_bef_hitplus_cat_FALSE", v)) return("glaucoma_bef_hitplus_cat_0")
    if (grepl("glaucoma_bef_hitplus_cat_TRUE", v)) return("glaucoma_bef_hitplus_cat_1")
    if (grepl("otherretina_bef_hitplus_cat_FALSE", v)) return("otherretina_bef_hitplus_cat_0")
    if (grepl("otherretina_bef_hitplus_cat_TRUE", v)) return("otherretina_bef_hitplus_cat_1")
    return(v)
  }

  # Standardize variable names
  original$variable_std <- sapply(original$variable, standardize_var)
  submitted$variable_std <- sapply(submitted$variable, standardize_var)
  excluded$variable_std <- sapply(excluded$variable, standardize_var)
  imputed$variable_std <- sapply(imputed$variable, standardize_var)

  # Get unique standardized variables
  vars <- unique(c(original$variable_std, submitted$variable_std, 
                   excluded$variable_std, imputed$variable_std))

  # Remove duplicates and unwanted rows
  vars <- vars[!grepl("^outcome_va_vi_binary$|^ever_lapse_binary$", vars)]

@@

    # Original
    orig_row <- original[original$variable_std == v, ]
    if (nrow(orig_row) > 0 && !is.na(orig_row$n)) {
      table_df$Original[i] <- sprintf("%.0f (%.1f%%)", orig_row$n, orig_row$pct)
    }

    # Submitted
    sub_row <- submitted[submitted$variable_std == v, ]
    if (nrow(sub_row) > 0 && !is.na(sub_row$n)) {
      table_df$Submitted[i] <- sprintf("%.0f (%.1f%%)", sub_row$n, sub_row$pct)
    }

    # Excluded
    exc_row <- excluded[excluded$variable_std == v, ]
    if (nrow(exc_row) > 0 && !is.na(exc_row$n)) {
      table_df$Excluded[i] <- sprintf("%.0f (%.1f%%)", exc_row$n, exc_row$pct)
    }

    # Imputed
    imp_row <- imputed[imputed$variable_std == v, ]
    if (nrow(imp_row) > 0 && !is.na(imp_row$n)) {
      table_df$Imputed[i] <- sprintf("%.1f (%.1f%%)", imp_row$n, imp_row$pct)
    }
  }

  # Clean up variable names for display
  table_df$Variable <- gsub("_0$", "_No", table_df$Variable)
  table_df$Variable <- gsub("_1$", "_Yes", table_df$Variable)
